---
title: 'Dimensionality Reduction'
author: 'Alekhya Pinnamaneni and Aloksai Choudari'
date: 'October 9, 2022'
output: 
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

### Select a data set
Data set: Metro Interstate Traffic Volume Data

Source: <https://archive.ics.uci.edu/ml/datasets/Metro+Interstate+Traffic+Volume>

Target column: 'traffic_volume'

No. of rows: 48,205 rows

### Load the data
```{r}
traffic <- read.csv("traffic.csv")
```

### Clean the data
```{r}
traffic$holiday <- as.numeric(as.integer(factor(traffic$holiday)))
traffic$temp <- as.numeric(as.integer(factor(traffic$temp)))
traffic$rain_1h <- as.numeric(as.integer(factor(traffic$rain_1h)))
traffic$snow_1h <- as.numeric(as.integer(factor(traffic$snow_1h)))
traffic$clouds_all <- as.numeric(as.integer(factor(traffic$clouds_all)))
traffic$weather_main <- as.numeric(as.integer(factor(traffic$weather_main)))
traffic$weather_description <- as.numeric(as.integer(factor(traffic$weather_description)))
traffic$date_time <- as.numeric(as.integer(factor(traffic$date_time)))
traffic$traffic_volume <- as.numeric(as.integer(factor(traffic$traffic_volume)))
```

### Split the data into train and test data
```{r}
set.seed(1234)
sample <- sample(1:nrow(traffic), nrow(traffic)*0.8, replace=FALSE)
train <- traffic[sample,]
test <- traffic[-sample,]
```

### PCA Dimensionality Reduction
```{r}
library(caret)

pca_out <- preProcess(train[,1:4], method=c("center", "scale", "pca"))
pca_out
```

### LDA Dimensionality Reduction

```{r}
library(MASS)
lda1 <- lda(temp~., data=train)
```

```{r}
lda_pred <- predict(lda1, newdata=test, type="class")
lda_pred$class
mean(lda_pred$class==test$temp)
```

### Analysis

Upon reduction of the data, a quite a bit of accuracy was lost. After PCA, the accuracy was similar as seen to the results in Notebook 1 - regression, but after LCA there was a significant decrease in accuracy. 